---
title: "Stores"
author: "Kevien Otieno"
date: "2024-02-17"
output: html_document
---

```{r}
#loading packages
pacman::p_load(tidyverse, tidyr, dplyr, ggplot2,lubridate)
```

```{r}
#loading our dataset
Stores=read.csv("Global Superstore.csv")
```

```{r}
#Lets View first 5 rows of our dataset
head(Stores, 5)
```

```{r}
#lets check on the dimensions of our dataset
dim(Stores)
```

```{r}
#list all column names
names(Stores)
```

```{r}
#lets check on the datatypes of our columns
str(Stores)
```

```{r}
#Separating the dates into other new columns i.e shipping day, month or year
Stores=Stores %>% 
  separate(col="Order.Date", 
           into=c("Order.Day", "Order.Month","Order.Year"),
           sep="[-/]", remove = FALSE)

Stores=Stores %>% 
  separate(col="Ship.Date",
           into=c("Ship.Day","Ship.Month","Ship.Year"),
           sep="[-/]", remove = FALSE)

```

```{r}
#converting a few columns into the necessary wanted datatypes
Stores$Order.Day=as.integer(Stores$Order.Day)
Stores$Order.Month=as.integer(Stores$Order.Month)
Stores$Order.Year=as.integer(Stores$Order.Year)
Stores$Ship.Day=as.integer(Stores$Ship.Day)
Stores$Ship.Month=as.integer(Stores$Order.Month)
Stores$Ship.Year=as.integer(Stores$Ship.Year)
Stores$Ship.Mode=as.factor(Stores$Ship.Mode)
Stores$Category=as.factor(Stores$Category)
Stores$Order.Priority =as.factor(Stores$Order.Priority)
Stores$Ship.Month=as.character(Stores$Order.Month)
Stores$Order.Month=as.character(Stores$Order.Month)
```


```{r}
#renaming the months into the expected
Stores=Stores %>% 
  mutate(
    Order.Month =case_when(Order.Month == 1 ~ "January",Order.Month == 2 ~ "February",
                         Order.Month == 3 ~ "March",Order.Month == 4 ~ "April",
                         Order.Month == 5 ~ "May",Order.Month == 6 ~ "June",
                         Order.Month == 7 ~ "July",Order.Month == 8 ~ "August",
                         Order.Month == 9 ~ "September",Order.Month == 10 ~ "October",
                         Order.Month == 11 ~ "November",Order.Month == 12 ~ "December"))

Stores=Stores %>% 
  mutate(
    Ship.Month =case_when(Ship.Month == 1 ~ "January",Ship.Month == 2 ~ "February",
                         Ship.Month == 3 ~ "March",Ship.Month == 4 ~ "April",
                         Ship.Month == 5 ~ "May",Ship.Month == 6 ~ "June",
                         Ship.Month == 7 ~ "July",Ship.Month == 8 ~ "August",
                         Ship.Month == 9 ~ "September",Ship.Month == 10 ~ "October",
                         Ship.Month == 11 ~ "November",Ship.Month == 12 ~ "December"))

```

```{r}
#checking the datatypes again
str(Stores)
```

Working on the date columns

```{r}
#lets now get the summary for our dataset
summary(Stores)
```
		Perform an Exploratory Data Analysis
 Customers Analysis		
1.Profile the customers based on their frequency of purchase - calculate frequency of purchase for each customer
```{r}
freq_purch=Stores %>% 
  select(Customer.ID, Customer.Name) %>% 
  group_by(Customer.Name) %>% 
  summarise(Frequency=n()) %>% 
  arrange(Frequency, by_group = TRUE)

Most_frequent_customers=tail(freq_purch,5)
```

2.Do the high frequent customers are contributing more revenue	
```{r}
Stores %>% 
  select(Customer.ID, Customer.Name, Sales) %>% 
  group_by(Customer.Name) %>%
  summarise(Total_sales=sum(Sales),Frequency=n()) %>% 
  arrange(desc(Total_sales),desc(Frequency)) %>% 
  head(5)

```
NO.The frequent customers do not contribute much to the revenue

3.Are they also profitable - what is the profit margin across the buckets		
```{r}
Stores %>% 
  select(Customer.ID, Customer.Name, Profit) %>% 
  group_by(Customer.Name) %>%
  summarise(Total_Profits=sum(Profit),Frequency=n()) %>% 
  arrange(desc(Total_Profits),desc(Frequency)) %>% 
  head(5)
```


4.Which customer segment is most profitable in each year.	
```{r}
segment_customer= Stores %>% 
  select(Segment, Profit) %>% 
  group_by(Segment) %>% 
  summarise(Total_Profits=ceiling(sum(Profit))) %>% 
  arrange(desc(Total_Profits))

pie(segment_customer$Total_Profits,
    labels =NA,
    lwd=2,
    main = "Profits per Segment",
    border="white",
    col = c("red","blue","green",
    cex.main =1.5
    ))

legend("topright", legend=segment_customer$Segment,fill = c("red","blue","green", title = "Segments"))
```

5.How are the customers distributed across the countries?		
```{r}
dist_Country=Stores %>% 
  select(Country) %>% 
  group_by(Country) %>% 
  summarise(Frequency=n())
```

Product Analysis		
1.Which country has top sales?	
```{r}
c_top=Stores %>%
  select(Country, Sales) %>% 
  group_by(Country) %>% 
  summarise(Total_Sales=sum(Sales)) %>% 
  arrange(desc(Total_Sales)) %>% 
  head(5)

ggplot(c_top, aes(x=Country, y=Total_Sales))+
  geom_bar(stat="identity",col="black",fill=c("red","yellow","green","blue","pink"))+
  labs(title = "Investigating the country with the most sales", x="Countries",y="Total Sales")+
  theme_minimal()+
  scale_fill_manual()+
  theme(plot.title = element_text(hjust=0.5))
  
```

2.Which are the top 5 profit-making product types on a yearly basis?	
```{r}
profit_product <- 
  rbind(
  (Stores %>% 
  select(Ship.Year, Sub.Category, Profit) %>% 
  filter(Ship.Year == 2011) %>% 
  group_by(Ship.Year,Sub.Category) %>% 
  summarise(Total_profits = ceiling(sum(Profit)), .groups = 'drop') %>% 
  arrange(desc(Total_profits)) %>% 
  head(5)),
  (Stores %>% 
  select(Ship.Year, Sub.Category, Profit) %>% 
  filter(Ship.Year == 2012) %>% 
  group_by(Ship.Year,Sub.Category) %>% 
  summarise(Total_profits = ceiling(sum(Profit)), .groups = 'drop') %>% 
  arrange(desc(Total_profits)) %>% 
  head(5)),
  (Stores %>% 
  select(Ship.Year, Sub.Category, Profit) %>% 
  filter(Ship.Year == 2013) %>% 
  group_by(Ship.Year,Sub.Category) %>% 
  summarise(Total_profits = ceiling(sum(Profit)), .groups = 'drop') %>% 
  arrange(desc(Total_profits)) %>% 
  head(5)),
  (Stores %>% 
  select(Ship.Year, Sub.Category, Profit) %>% 
  filter(Ship.Year == 2014) %>% 
  group_by(Ship.Year,Sub.Category) %>% 
  summarise(Total_profits = ceiling(sum(Profit)), .groups = 'drop') %>% 
  arrange(desc(Total_profits)) %>% 
  head(5)),
  (Stores %>% 
  select(Ship.Year, Sub.Category, Profit) %>% 
  filter(Ship.Year == 2015) %>% 
  group_by(Ship.Year,Sub.Category) %>% 
  summarise(Total_profits = ceiling(sum(Profit)), .groups = 'drop') %>% 
  arrange(desc(Total_profits)) %>% 
  head(5)))
  
ggplot(profit_product, aes(x=Ship.Year, y=Total_profits, fill=Sub.Category))+
  geom_bar(stat = "identity", position = "dodge")+
  labs(title = "Total Profits per year and product", x= "Year", y= "Total Profits")+
  theme_minimal()+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_fill_manual(values = c("red","blue","green","black","purple","orange","brown"))

```


3.How is the product price varying with sales - Is there any increase in sales with the decrease in price at any level?
```{r}

```

4.What is the average delivery time across the countries?

5.What is the profit by month? (Use a line chart to visualize this.)	
```{r}
profit_month=Stores %>% 
  select(Order.Month, Profit) %>% 
  group_by(Order.Month) %>% 
  summarise(Total_Profits=ceiling(sum(Profit)))

profit_month$Order.Month <- factor(profit_month$Order.Month, levels = month.name)

ggplot(profit_month, aes(x = Order.Month, y = Total_Profits, group=1)) +
  geom_line(color = "blue", linewidth = 1.5) + 
  labs(title = "Total Profit by Month", x = "Month", y = "Total Profit") +
  theme_minimal() + 
  theme(
    plot.title = element_text(hjust = 0.5),  
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```
6.Profit by category? By Subcategory?		
```{r}
profit_category=Stores %>% 
  select(Category, Profit) %>% 
  group_by(Category) %>% 
  summarise(Total_Profits=ceiling(sum(Profit)))

profit_subcategory=Stores %>% 
  select(Sub.Category, Profit) %>% 
  group_by(Sub.Category) %>% 
  summarise(Total_Profits=ceiling(sum(Profit)))

print(profit_subcategory)


pie(profit_category$Total_Profits, 
    labels = profit_category$Category,
    lwd=2,
    main="Profits per Category",
    col=c("red", "blue","pink"),
    border = "white",  # Add white border around each slice
  lwd = 2,  # Increase line width for slice outlines
  clockwise = TRUE,  # Change the direction of the slices to clockwise
  init.angle = 90,  # Start drawing slices from the top (90 degrees)
  cex.main = 1.5,  # Increase the size of the main title
  cex.lab = 1.2,   # Increase the size of the labels
  cex.axis = 0.8   # Adjust the size of the axis text
  )

legend("topright", legend=profit_category$Category, fill=c("red","blue","green"), title="Categories")
```
7.Sales by month? By category? By subcategory?		

```{r}
sales_month=Stores %>% 
  select(Order.Month, Sales) %>% 
  group_by(Order.Month) %>% 
  summarise(Total_Sales=ceiling(sum(Sales)))

sales_month$Order.Month <- factor(sales_month$Order.Month, levels = month.name)

sales_category=Stores %>% 
  select(Category, Sales) %>% 
  group_by(Category) %>% 
  summarise(Total_Sales=ceiling(sum(Sales)))

sales_subcategory=Stores %>% 
  select(Sub.Category, Sales) %>% 
  group_by(Sub.Category) %>% 
  summarise(Total_Sales=ceiling(sum(Sales)))

print(sales_subcategory)

pie(sales_category$Total_Sales,
    labels = NA,
    lwd=2,
    main="Sales per Category",
    col=c("red","blue","green"),
    border = "white",
    cex.main=1.5
    )

legend("topright", legend = sales_category$Category, fill = c("red","blue","green"), title = "Categories")

ggplot(sales_month, aes(x=Order.Month, y= Total_Sales))+
  geom_bar(stat="identity", col="black", fill="blue")+
  labs(title = "Total Sales per Month", x= "Months", y="Total Sales")+
  theme_minimal()+
  theme(
    plot.title = element_text(hjust=0.5),
    axis.text.x = element_text(angle = 45, hjust=1)
  )
```
